---
title: "PA1_template"
author: "agrainofsand"
output: html_document
---

## Loading and preprocessing the data
```{r,echo=TRUE, message=FALSE}
# Loading library
library(ggplot2)
library(dplyr)
library(gridExtra)
activity <- read.csv("/Users/agrainofsand/Documents/coursera_datascience/5 Reproducible Research/activity.csv",
                     stringsAsFactors=F)
```


## What is mean total number of steps taken per day?
```{r,echo=TRUE}
activity$date <- as.Date(activity$date)
steps_by_date <- aggregate(data=activity, steps ~ date, FUN=sum)

print('1. Total number of steps per day:')
steps_by_date

print('2. Histogram of total number of steps per day')
hist(steps_by_date$steps, main = "Histogram of Total Number of Steps per Day", xlab="Steps")

mean_steps <- mean(steps_by_date$steps)
paste('3a. Mean steps: ', round(mean_steps,0))

median_steps <- median(steps_by_date$steps)
paste('3b. Median steps: ', median_steps)
```
Mean total of steps taken per day is `r round(mean_steps,2)`


## What is the average daily activity pattern?
1. Average daily pattern
```{r, echo=TRUE}
avg_daily_pattern <- aggregate(activity$steps, by = list(activity$interval), FUN=mean, na.rm = T)
colnames(avg_daily_pattern) <- c("intervals", "mean")

plot(avg_daily_pattern$intervals, avg_daily_pattern$mean, xlab = "Time interval of 5 min", ylab = "Avg number of steps 5 min interval", type = "l", main = "Avg daily activity pattern")

```

2. Find interval with max steps
```{r,echo=TRUE}
max_steps <- max(avg_daily_pattern$mean)
interval_with_max_steps <- avg_daily_pattern[avg_daily_pattern$mean==max(max_steps),1]
interval_with_max_steps
```


## Imputing missing values
1. Calculating and reporting total number of missing values
```{r, echo=TRUE}
number_of_rows_with_missing_values <- nrow(filter(activity, is.na(steps)))
number_of_rows_with_missing_values
```

2. Create a new data set, filling in missing values, using the mean of each interval
```{r, echo=TRUE}
activity_new <- merge(activity, avg_daily_pattern, by.x="interval", by.y="intervals") %>%
                  mutate(steps_new = ifelse(is.na(steps),mean, steps)) %>%
                  select(steps_new, date, interval)

```

3. Histogram of total number of steps taken each day
```{r, echo=TRUE}
activity_new_by_date <- aggregate(steps_new ~ date, activity_new, sum)

hist(activity_new_by_date$steps, xlab ="Steps", ylab = "Frequency", main="New Number of steps taken per day")

```

4. Mean and median of total number of steps now
``` {r, echo=TRUE}
mean_steps_new <- mean(activity_new_by_date$steps)
median_steps_new <- median(activity_new_by_date$steps)
```
Mean steps now is `r round(mean_steps_new,2)`, which is the same as the mean number of steps before, given that we used the mean to fill in the NAs.

Median steps now is `r round(median_steps_new,2)`, which is `round(median_steps_new - median_steps,2)` different from the original median step count of `r round(median_steps,2)`.

## Are there differences in activity patterns between weekdays and weekends?
``` {r, echo=TRUE}
activity_new_by_date <- activity_new_by_date %>%
                          mutate(weekday = weekdays(as.Date(date)),
                                  is_weekend = ifelse(weekday == "Saturday" ,"yes", ifelse(weekday == "Sunday","yes","no"))
                                  )

mean_steps_weekend <- aggregate(steps_new ~ is_weekend, data=activity_new_by_date, mean)
mean_steps_weekend
```
1. The mean number of steps on weekends is `r round(filter(mean_steps_weekend, is_weekend=="yes")[,2],2)`, which is `r round((filter(mean_steps_weekend, is_weekend=="yes")[,2] / filter(mean_steps_weekend, is_weekend=="no")[,2] -1),2) *100`% higher than the mean number of steps on weekdays.

2. Panel plot
``` {r, echo=TRUE}
activity_new_by_interval <- activity_new %>%
                          mutate(weekday = weekdays(as.Date(date)),
                                  is_weekend = ifelse(weekday == "Saturday" ,"yes", ifelse(weekday == "Sunday","yes","no"))
                                  )
activity_new_by_interval_mean <- aggregate(data = activity_new_by_interval, steps_new ~ interval + is_weekend, FUN=mean)
weekday_activity <- filter(activity_new_by_interval_mean, is_weekend=="no")
weekend_activity <- filter(activity_new_by_interval_mean, is_weekend=="yes")

graph.weekday_activity <- ggplot(data=weekday_activity, aes(x=interval, y=steps_new)) + geom_line() + theme_bw() + ggtitle("Weekday Activity")

graph.weekend_activity <- ggplot(data=weekend_activity, aes(x=interval, y=steps_new)) + geom_line() + theme_bw() + ggtitle("Weekend Activity")

grid.arrange(graph.weekday_activity, graph.weekend_activity, nrow=2)
```
